---
title: "Preprocess & Explore iPhone Small Matrix"
author: "Katherine Piatti"
date: "4/29/2021"
output: 
 html_document:
 theme: united
 toc: TRUE
 toc_depth: 2
 toc_float: TRUE
 highlight: tango
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# SETUP ##########################################
  
## Project Overview ##############################

The goal of this project is to use text mining to gather data on sentiments towards iPhone and Samsung Galaxy mobile devices from archived webpages from the CommonCrawl archive.

In this notebook I will be analyzing and modeling the iphone data from the large matrix of text mining data I gathered from a portion of the February 2021 CommonCrawl database. 


## Packages & Renv #####################################

  - I ran renv::init 2021-39-04 2:08PM
  
```{r}
#load pkgs
library(tidyverse)
library(here)
library(janitor)
library(corrr)
library(GGally)
library(stargazer)
library(lattice)

library(broom)
library(caret)
library(MLmetrics)
library(doParallel)
```

## Import Data #######################################

```{r}
#read in iphone small matrix
lgiphone_df_orig <- read.csv(here("data", "LargeMatrix.csv")) 

# define copy of original data, clean var names, & remove any empty columns or rows
lgiphone_df <- lgiphone_df_orig %>%
  clean_names() %>% 
  remove_empty()
```