---
title: "Preprocess & Explore iPhone Small Matrix"
author: "Katherine Piatti"
date: "4/29/2021"
output: 
 html_document:
 theme: united
 toc: TRUE
 toc_depth: 2
 toc_float: TRUE
 highlight: tango
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

options(scipen = 1, digits = 3) #set to 3 decimal 
```

# SETUP ##########################################
  
## Project Overview ##############################

The goal of this project is to use text mining to gather data on sentiments towards iPhone and Samsung Galaxy mobile devices from archived webpages from the CommonCrawl archive.

In this notebook I will be analyzing and modeling the iphone data from the large matrix of text mining data I gathered from a portion of the February 2021 CommonCrawl database. 


## Packages & Renv #####################################

  - I ran renv::init 2021-39-04 2:08PM
  
```{r}
#load pkgs
library(tidyverse)
library(here)
library(janitor)
library(corrr)
library(GGally)
library(stargazer)
library(lattice)

library(broom)
library(caret)
library(MLmetrics)
library(doParallel)
```

## Import Data #######################################

```{r}
#read in iphone small matrix
lgiphone_df_orig <- read.csv(here("data", "LargeMatrix.csv")) 

# define copy of original data, clean var names, & remove any empty columns or rows
lgiphone_df <- lgiphone_df_orig %>%
  clean_names() %>% 
  remove_empty()

#verify import
glimpse(lgiphone_df)
```


# DATA WRANGING #########

The raw data has 30,701 rows and 59 columns.


## Remove Irrelevant Variables #######

```{r}
#select only columns that start with iphone or ios
lgiphone_df <- lgiphone_df %>% 
  select(starts_with(c("iphone", "ios"))) 

glimpse(lgiphone_df)
```

- The data now has 30,701 rows and 14 columns



## Missing Values #####################

```{r}
#check for missing values
lgiphone_df %>% 
  is.na() %>% 
  sum()
```

```{r}
# return all unique values in each var
lgiphone_df %>% 
  map(unique)
```



## Duplicates #############

```{r}
#get total number of duplicates
lgiphone_df %>% 
  duplicated() %>% 
  sum()
```


```{r}
#keep only distinct rows
lgiphone_df %>% 
  distinct() %>% 
  glimpse()

lgiphone_distinct <- lgiphone_df %>% 
  distinct()
```



## Feature Correlations ############

```{r}
#get var correlations > 0.2 and -0.2 for iphone_df
(corr <- lgiphone_df %>% 
  correlate() %>% 
  shave() %>% 
  stretch() %>% 
  filter(abs(r) > 0.2) %>% 
  arrange(desc(r)))
```
_Observations_
- It is notable that the high correlations between several iphone variables that we saw in the small iphone matrix are not present in this dataset. This is reason to doubt that the any model trained on the small matrix could predict overall iphone sentiment on this data. 


```{r}
# plot bivariate relationships & densities for display vars 
lgiphone_df %>% 
  select(c(iphonedispos, iphonedisneg, iphonedisunc)) %>% 
  ggpairs()
```

```{r}
# plot bivariate relationships & densities for performance vars
lgiphone_df %>% 
  select(iphoneperpos, iphoneperneg, iphoneperunc) %>% 
  ggpairs()
```

_Observations_
- There is something profoundly wrong with this data. 

```{r}
#get var correlations > 0.2 and -0.2 for iphone_df
(corr <- lgiphone_distinct %>% 
  correlate() %>% 
  shave() %>% 
  stretch() %>% 
  filter(abs(r) > 0.2) %>% 
  arrange(desc(r)))
```




## Feature Variances ###############

*Data with Duplicates*
```{r}
#look for vars with near zero variance
lgiphone_df %>% 
  nearZeroVar(saveMetrics = TRUE) %>% 
  filter(nzv == "TRUE")
```

*Data without Duplicates*
```{r}
#look for vars with near zero variance
lgiphone_distinct %>% 
  nearZeroVar(saveMetrics = TRUE) %>% 
  filter(nzv == "TRUE")
```




## Remove iOS Sentiment Variables ##################

```{r}
#remove iOS sentiment vars 
lgiphone_df <- lgiphone_df %>% 
  select(-c(iosperpos, iosperneg, iosperunc)) %>%
  glimpse()
```

```{r}
#remove iOS sentiment vars
lgiphone_df <- lgiphone_distinct %>% 
  select(-c(iosperpos, iosperneg, iosperunc)) %>%
  glimpse()
```




# DATA EXPLORATION ###############

## Summary Stats ############

```{r}
stargazer(lgiphone_df, type = "text")
```

```{r}
stargazer(lgiphone_distinct, type = "text")
```



## Histograms ################

```{r}
#create histograms of all vars in df with dupes
lgiphone_df %>% 
  map(hist)
```

```{r}
#create histograms of all vars in df with dupes
lgiphone_distinct %>% 
  map(hist)
```



## Add Target Column ###############

Because the overall sentiment towards iphones was not hand coded for this dataset, I need to add a "iphonesentiment" column to hold the model predictions

```{r}
#add iphonesentiment column to end of df
lgiphone_df <- lgiphone_df %>% 
  add_column(iphonesentiment = NA)

#verify column added
glimpse(lgiphone_df)
```

```{r}
#add iphonesentiment column to end of df
lgiphone_distinct <- lgiphone_df %>% 
  add_column(iphonesentiment = NA)

#verify column added
glimpse(lgiphone_df)
```

>*Coding Note*
>
- You can add empty columns before or after other columns by adding .before = varname or .after = varname parameter to add_column func.



# Parallel Processing ###########################

```{r}
#get numberr of cores on my machine
detectCores()

# create cluster
ppcluster <- makeCluster(3)

#register cluster
registerDoParallel(ppcluster)

# get number of cores now assigned to R
getDoParWorkers()
```




# PREDICT OVERALL SENTIMENT #####################################

```{r}
# use random rf_fit01.1 to predict iphone sentiment
lgiphone_df_rfpreds <- predict(rf_fit02.1, lgiphone_df)

#plot distribution of predicted sentiment
rfpreds <- barchart(lgiphone_df_rfpreds,
                    main = "rfpreds")
```

```{r}
# use random rf_fit01.1 to predict iphone sentiment
lgiphone_df_c50preds <- predict(c50_fit02.1, lgiphone_df)

#plot distribution of predicted sentiment
c50preds <- barchart(lgiphone_df_c50preds,
                     main = "c50preds")
```

```{r}
print(rfpreds, split = c(1,1,2,1), more = TRUE)
print(c50preds, split = c(2,1,2,1), more = FALSE)
```


# SUMMARY ##################

- Given how unusual characteristics of the data in my large matrix and how different it is from the data in the small iphone matrix, I do not think the predictions are trustworthy. 
